---
title: "final project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(stringr)
library(gridExtra)
library(stargazer)
library(readr)
library(foreign)
library(lfe)
library(tidyr)
library(reshape)
library(ebal)

water_df <- read.csv("/Users/jiahuizhang/Documents/UCLA 2.2/Causal inferences/Final Project/water.csv")

```

```{r}
# wide to long data
# ID might need to be factor
watero_df <- water_df %>% 
  select(id, age, height, caloric_intake, pre_sleep, pre_exercise, pre_weight, pre_composition, pre_water, Di, drop_out) %>% 
  mutate(postperiod = 0)

watero_df <- rename(watero_df, c(pre_weight="weight", pre_composition = "composition", pre_water = "water" ))

water1_df <- water_df %>% 
  select(id, age, height, caloric_intake, pre_sleep, pre_exercise, post_weight, post_composition, post_water, Di, drop_out) %>% 
  mutate(postperiod = 1)

water1_df <- rename(water1_df, c(post_weight="weight", post_composition = "composition", post_water = "water" ))

waterdata_df <- rbind(watero_df, water1_df)

write.csv(waterdata_df, file = "/Users/jiahuizhang/Documents/UCLA 2.2/Causal inferences/Final Project/waterdata.csv")
```

1. Non-compliance, whether some girls from the treatment group were not treated, whether some girls from the control group received treatment

```{r}
treatment <- water_df %>%
  mutate(extrawater = post_water - pre_water) %>%
  filter(Di == 1) 
shapiro.test(treatment$extrawater)
require(graphics)
qqnorm(treatment$extrawater);qqline(treatment$extrawater, col = 2)
t.test(treatment$extrawater, mu=1.5)
# shapiro test suggests it's normally distributed; ttest can be interpreted as there's no non-compliance for the treatment group

control <- water_df %>%
  mutate(extrawater = post_water - pre_water) %>%
  filter(Di == 0) 
shapiro.test(control$extrawater)
require(graphics)
qqnorm(control$extrawater);qqline(control$extrawater, col = 2)
t.test(control$extrawater, mu=0)
# shapiro test suggests it's normally distributed; ttest can be interpreted as there's no non-compliance for the control group
  
```
The null-hypothesis of the shapiro test is that the population is normally distributed. Thus, if the p-value is less than the chosen alpha level, then the null hypothesis is rejected and there is evidence that the data tested are not from a normally distributed population; in other words, the data are not normal.

Conclusion: only the treatment group received treatment and all of them received treatment; no one in the control group received any treatment and the amount of water they drunk did not differ significantly from what they had drunk before the experiment.  

2. Balance table: first balance table for two groups before treatment; then balance table for two groups considering sample attrition.
```{r}
water_df <- water_df %>% 
  filter(drop_out == 0) %>%
  mutate(treatment = relevel(factor(Di), ref = "0"))


# Estimating differences between groups:
m_list <- list()
m_list[[1]] <- felm(age ~ treatment | 0 | 0 | id, data = water_df)
m_list[[2]] <- felm(height ~ treatment | 0 | 0 | id, data = water_df) 
m_list[[3]] <- felm(caloric_intake ~ treatment | 0 | 0 | id, data = water_df)


# Creating Table:
stargazer(m_list,
type = "text",
digits = 2,
intercept.bottom = FALSE,
dep.var.labels = c("age", "height", "caloric_intake"), 
covariate.labels = c("Control Mean", "treatment"), model.numbers = FALSE,
report = "vcp*",
star.cutoffs = c(.05, NA, NA),
notes = "$^{*}$p$<$0.05; Clustered SEs",
keep.stat = "n",
title = "Balance by Pre-Treatment Covariates",
no.space = TRUE)

```

```{r}
m_list2 <- list()
m_list2[[1]] <- felm(pre_sleep ~ treatment | 0 | 0 | id, data = water_df) 
m_list2[[2]] <- felm(pre_exercise ~ treatment | 0 | 0 | id, data = water_df) 
m_list2[[3]] <- felm(pre_water ~ treatment | 0 | 0 | id, data = water_df) 

stargazer(m_list2,
type = "text",
digits = 2,
intercept.bottom = FALSE,
dep.var.labels = c("sleep", "exercise", "water"), 
covariate.labels = c("Control Mean", "treatment"), 
model.numbers = FALSE,
report = "vcp*",
star.cutoffs = c(.05, NA, NA),
notes = "$^{*}$p$<$0.05; Clustered SEs",
keep.stat = "n",
title = "Balance by Pre-Treatment Covariates",
no.space = TRUE)


```

```{r}
m_list[[7]] <- felm(pre_weight ~ treatment | 0 | 0 | id, data = water_df) 
m_list[[8]] <- felm(pre_composition ~ treatment | 0 | 0 | id, data = water_df) 

```



```

 




